{\normalsize Seminar: Formal Specification} \\[1ex] Security of Multithreaded Programs by Compilation\cite{Barthe07}

### 2012-01-29 paper/paper.tex
revied whole paper

\section{Motivation}
\label{sec:motivation}
Protecting the confidentiality of information that is processed by
a modern computer is challenging and important, since more and more sensitive
information is fed to them. Access control and encryption are not enough to
ensure this confidentiality, because the usage of the information after the
access or decryption is not restricted. It is necessary to control the
\textit{information flow} to protect the sensitive information.
On modern mobile devices sensitive information is processed and programs
are executed multithreaded (e.g. to prevent lock ups when establishing
a network connection). Through the timing difference of the scheduled threads it
is likely that sensitive information is leaked and an attack may obtain all
sensitive data.

The idea Barthe, Rezk, Russo and Sabelfeld phrased in \cite{Barthe07} was to close
this leak (more formally called \textit{covert channel}) by annotating the byte-code
of the program (most mobile devices use some sort of byte-code e.g. Android) with security
labels according to the processed information, which causes the scheduler to hide
threads that leak information from the attacker.

\section{Introduction to the research area}
\label{sec:introduction}
The challenge to protect the confidentiality of information processed by
a computer lead thirty years ago (c.f. \cite{Zdancewic04}) to the research
on the security information-flows, because of the shortcomings of the standard methods.
These standard methods are: Access control and encryption. Where
access control ensures that the data is only accessed by an authorized entity,
and encryption ensures that data can be securely transmitted over an insecure channel.
But all these methods cover the release of information and not the further
propagation (i.e. how the information is used after having access to it or decrypting
it). This is the place where information-flow security comes into play.

The idea is to \textit{"track and regulate"}\cite{Zdancewic04} where information flows to
prevent the leak of secret information. The ways on which the information flows
through the system are called channels. Channels that are not indented to transport
information are called covered channels. These covered channels can
be classified into the following categories \cite{Sabelfeld03} (in the following examples $h$ is a secret
variable and $l$ a variable that is plublic):

\begin{itemize}
\item \textit{Explicit and implicit flows}: In an explicit flow the information is leaked
      directly into some public variable (e.g. $l := h$), while in implicit flows the information
      is leaked through the control structure of the program (e.g. $\text{if}\ h = 1\ \text{then}\ l := 1\ \text{else}\ l := 0$).
\item \textit{Probabilistic channels}: If an attacker is able to run a computation multiple times
      he might be able to obtain information by looking at the probability distribution of the
      public outputs.
\item \textit{Power channels}: If the attacker has physical access to the computer or at least figure out
      the power consumption, he might be able to obtain data through the changing power consumption.
\item \textit{Resource exhausting channels}: Information also may be leaked through the exhaustion
      of finite (mostly physical) resources (e.g. a buffer overflow).
\item \textit{Termination channels}: The attacker can obtain information through the termination
      or non-termination of the computation or program.
\item \textit{Timing channels:} Information can be obtained through the time at which this action
      occurs. On the one hand \textit{external} timing channels cover the action like the termination
      of the program, where the attacker obtains information from the total execution time of the program.
      On the other hand \textit{internal} timing channels are occur in multithreaded programs through the timing difference
      between threads. For example the two threaded (denoted with $||$) program \[\text{if}\ h = 1\ \{\text{sleep}(100)\}\ ;\ l := 1\ ||\ \text{sleep}(50)\ ;\ l := 0\]
      executed with most schedulers will leak the information of $h$ into $l$ through the timing difference
      if $h = 1$.
\end{itemize}

The notion of confidentiality that is used in most work in the area of information-flow security is \textit{noninterference}.
This policy is defined in various ways but the essence is, that it requires \textit{"that secret information
[do] not affect [the] publicly observable behavior of a system"}\footnote{Even wen the attacker has full access to the
source code of the program.}\cite{Zdancewic04}. For most real world
applications this policy is far too strict, because it forbids useful programs like password checkers.\footnote{A
password checker needs to reveal whether the user input was correct or not.} Approaches that allow
controlled release of information are called declassification. The kinds of declassification are \textit{what}
information is released, \textit{who} released this information, \textit{where} in the system the information
is released and \textit{when} the information is released (c.f. \cite{Sabelfeld05}). For these approaches it is necessary that an
active attacker can only know as much as a passive attacker.

Mechanisms for controlling the information flow can be implemented either dynamic or static. The dynamic approach\footnote{
The taint mode of the programming language Perl uses this mechanism.}
labels the information with security labels and propagates these labels wherever the information is used. The static
approach\footnote{This approach is implemented in Jif for Java and Flow Caml for Caml.} analyses the program code and is
therefor far more promising, because with this approach it is possible to check all evaluation paths.

@@

\section{Summary of the article}
\label{sec:discussion}
In this summary I will follow mainly the structure of the original
paper \cite{Barthe07} which is as follows. After the introduction the basic
terms and definitions for a multithreaded programs and the scheduler are
laid. After that the notion of security we want to achieve is presented.
Along with this a skeleton of a type system is described, which ensures
that a program typable in this type system is secure w.r.t the notion of
security. The proof that this holds is sketched in the following section.
In the last section the (by now) abstract framework is instantiated with a
concrete example.

@@

These instructions are not further specified, but contain an instruction
to create a new thread (\texttt{start \textit{pc}} where pc is the start
instruction of the new thread).

@@

The next thing introduced are the security levels. We assume the attacker
"is" a level $k$. From this assumption we can reduce every set of security levels
w.l.o.g into $\{low, high\}$, where $low < high$, by mapping elements
that are no more sensitive than $k$ to $low$ and all other elements
-- including incomparable ones -- to $high$. It is also assumed, that access
control works correctly (i.e. the attacker can not access $high$ elements
directly).

@@

At this point a first simplification can be made. Looking a state $s \in \text{ConcState}$ we can
first extract the active threads (\texttt{s.act}) by taking the domain of the first
component. According to a security environment we can classify these threads,
with respect to their current program point (\texttt{s.pc(tid)} where $tid \in \text{Thread}$),
into \textit{low} threads iff the current program point is low. In \textit{high}
threads iff the current program point is high. In \textit{always high} threads
iff the current program point is always high and in \textit{hidden} threads, iff
the current program point is high but not always high.

The latter two are the interesting ones. If a thread is \textit{always high} it
can not leaks any information into low, because it never gets in touch
with low program points. So these threads can safely be interleaved between any
other threads by the scheduler.

@@

At this point no concrete scheduler is defined to make the framework applicable
for a wide class of schedulers. A scheduler is in this class iff it can be
modeled as a function $pickt: \text{ConcState} \times \text{History} \rightharpoonup \text{Thread}$
which satisfies the following constraints:

\begin{enumerate}
\item It always picks an active thread
\item If there is a hidden thread, always choose high or always high threads
\item Only use low information and the low part of the history to choose a
      new thread
\end{enumerate}

In constraint two the interleaving of always high threads is realized.

To define what noninterference means we need a notion of excution. It is assumed
that we have a sequential execution relation $\leadsto_{seq} \subseteq \text{SeqState} \times
\text{SeqState}$. The one step execution for the multithreaded language $\leadsto_{seq}
\subseteq (\text{ConcState} \times \text{History}) \times (\text{ConcState} \times \text{History})$ is defined
by the rules in figure \ref{fig:multithreaded-execution}.

@@

The scheduler is allow to pick a new thread after every transition of $\leadsto_{seq}$. The
intuition of the first rule is, that the scheduler picks a new thread $ctid$ which is at
program point $i$. This program point maps to an sequential instruction. The exection of
this instruction leads to a local state $\sigma$ and a global memory $\mu$. If these conditions
are met and instruction $i$ was not the last instruction (i.e. $\sigma.pc \neq \mathtt{exit}$),
then the concurent transition can be made. In this transition the concurrent state is updated
according to the results of the sequential execution and the thread id and security level of $i$
is recorded in the history.

The second rule covers the case in which the thread pickted by the scheduler has only the current
instruction before terminating (i.e. $\sigma.pc = \mathtt{exit}$). In the concurrent transition
the current thread id is removed from the concurrent state and everything else is like in the
first rule.

The third rule introduces the dynamic creation of new threads. In this rule the function $fresh_{l}$
\footnote{It is assumed that $fresh_l(tid) \neq fresh_{l'}(tid) \Leftrightarrow l \neq l'$.}
takes a set of thread identifiers and returns a new thread identifier at level $l$ and $\lambda_{init} : \mathcal{P}
\rightarrow \text{LocState}$ produces an initial state with the program pointer at the given program point.
If the current instruction at program point $i$ is the \texttt{start} instruction, a new thread
identifier is picked w.r.t. the security level of $i$. The program counter of the current thread
is increased by one, to step to the next instruction. The resulting concurrent state includes the
updated state of the current thread and the initial state for the new thread. The global memory
is not modified and the history is extended with the current thread id and the security level of $i$.

@@

So the state $s$ evaluates according to a history $h$ to a final global memory $\mu$ iff there is a
sequence of concurrent executions that terminates (i.e. there are no active threads) and has the
global memory $\mu$. $\leadsto_{conc}^*$ is the reflexive and transitive closure of of $\leadsto_{conc}$.

@@

Now we can define our goal: Noninterference. We define noninterference
in accordance to an indistinguishability relation $\sim_g$ on global memories.
Barthe et.al. state that it is not necessary, for the purpose of the paper,
to specify the definition of this relation. But to get a feel for this
relation, you can define it as: $\mu \sim_g \mu' \Leftrightarrow \mu|_{low} =
\mu'|_{low}$ where $\mu|_{low}$ projects out all high elements. Based on that
a program P is noninterfering if for all gobal memories $\mu_1, \mu_2, \mu_1'$
and $\mu_2'$ it holds that:

\begin{align*}
\mu_1 \sim_g \mu_2 \land P,\mu_1 \Downarrow_{conc} \mu_1' \land P,\mu_2 \Downarrow_{conc} \mu_2' \Rightarrow \mu_1' \sim_g \mu_2'
\end{align*}

\subsection{Type system}
\label{sec:typesystem}
The type system is the core part of the framework in the effect that it
enforces the previously defined noninterference property. Thus every
program which is typable, is noninterfering.

@@

The intuition of the typing judgment is that if we execute the instruction at program
point $i$ w.r.t. the security environment $se$ and the current type is $\textsc{s}$. Then the
type after the execution is $\textsc{t}$.

@@

A program is typeable in this type system (written $\mathcal{S}, se \vdash P$),
where $\mathcal{S}$ is a function $\mathcal{S}: \mathcal{P} \rightarrow \text{LType}$
that maps a local type to every program point
and $se$ a security environment, iff $\mathcal{S}$ maps every initial program point
(including the ones of spawned threads) to $\textsc{t}_{init}$ and for every program point $j$ which is an
successor of $i$ there is a type $s \in \text{LType}$, such that $\textsc{s}$ is lower bounded
by $\mathcal{S}(j)$ and $se, i \vdash \mathcal{S}(i) \Rightarrow \textsc{s}$ holds.

@@

\begin{theorem}
If the scheduler is secure and $se, \mathcal{S} \vdash P$, then P is noninterfering.
\end{theorem}

@@

The domain of next (i.e. NePd) captures the high, but not always high program points (i.e. the ones
that can result in a hidden thread). The property NeP1, states that two directly successive high
program points have the same program point in which the thread becomes visible again. The property NeP2
is the counterpart of the NeP1: If a low program point follows a high program point, this low program
point is the result of the next function for the high program point. NeP3 and NeP4 denote that the
next of an if respective while instruction is at least after the control dependence region.

@@

where $op \in \{+,-,\times,/\}$, $n \in \mathbb{Z}$ and $x$ are variables. The
operational semantics are standard and not explicitly necessary for the following
instantiation, therefor they are omited.

@@

Besides this concrete language we need to define a type system to enforce noninterference
according to section \ref{sec:typesystem}. The local types are defined by a stack of security levels
$\text{LType} = \text{Stack}(\text{Level})$ and $\textsc{t}_{init}$ with the empty stack. The typing rules defined in
figure~\ref{fig:multithreaded-typing-rules} need to be extended with rules for the concrete instructions
of the assembly language. In this summary I only explain two of them, the rest follows
in a similar manner.

@@

The security levels in $LType$ are the security levels of the operands in the local state. To
express our security policy, we declare a function $\Gamma(x)$ that assigns to every variable
$x$ a security level. If we now want to \texttt{store} the top of the stack into the variable
$x$ one premise is that the security level of the program point ($se(i)$) \textit{and} the
security level of the value on top of the operand stack is \textit{lower or equal} to the
security level that was assigned by $\Gamma$ to the variable $x$. If this is the case the head
of $LType$ is removed. The $\sqcup$ operator is like a logical "or", it returns the higher of
the given security values.

@@

Now we can define $tregion$ according to a compilation function $\mathcal{C}$. I won't define
the compilation function here, because it is not that essential how it exactly looks like.

@@

\section{Comparison with other approaches/further work}
\label{sec:furtherwork}
In this section I will first give a short overview of two other approaches and then
I will compare them with the approach presented before and draw a conclusion.

@@

They introduced the language $\lambda_{SEC}^{PAR}$ which supports higher-order functions, an
unbounded number of threads, synchronization (via join patterns), message passing and shared memory. Regardless of those realistic
features $\lambda_{SEC}^{PAR}$ is not intended to server as a user-lever programming language because its
syntax and type system is too awkward, instead it is a model for studying information flow.

Low-security observational determinism is used to ensure noninterference. The idea is
that to make the naturally nondeterministic system, deterministic from the point of view
of the attacker.

@@

\subsection{Flexible Scheduler-Independent Security\cite{Mantel10}}
This approach developed by Mantel and Sudbrock has the goal to be on the one hand
scheduler independent and on the other not too restrictive (i.e. does not treat
obviously secure programs as insecure).

To reach this goal they model the scheduling explicit but general enough to apply
to a wide class of common schedulers. So rather than talking about program states,
they introduces system configurations which contain the list of threads (thread pool) the global
memory and the scheduler state. The scheduler is able to store information in its
state and can retrieve input (e.g. the number of active threads). Based on this
information the scheduler makes decisions. To cover nondeterministic schedulers
(e.g. the uniform scheduler) the decision making of the scheduler and the execution
of the program is probabilistic.

Based on this scheduler model, a schedule-specific security property $\mathcal{S}$ is
defined on thread pools. Intuitively a thread pool is $\mathcal{S}$-secure if the
probability of running a program with system configuration $conf$ under the scheduler $\mathcal{S}$
results in the same global memory modulo low-equality as running the program with
system configuration $conf'$.

@@

In the end Barthe et. al. presented a framework with reaches many goals (like permissiveness, statically check-able
and requires no intervention from the programmer) but needs to modify the scheduler interface, which makes it
hard to use for real programs.

>>>>

\section{Motivation}
\label{sec:motivation}
Protecting the confidentiality of information that is processed by
a modern computer is challenging and important, since more and more sensitive
information is fed to them. Access control and encryption are not enough to
ensure this confidentiality, because the usage of the information after the
access or decryption is not restricted. It is necessary to control the
\textit{information flow} to protect the sensitive information.
On modern mobile devices sensitive information is processed and programs
are executed multithreaded (e.g. to prevent lock ups when establishing
a network connection). Through the timing difference of the scheduled threads it
is likely that sensitive information is leaked and an attacker may obtain
sensitive data.

The idea Barthe, Rezk, Russo and Sabelfeld phrased in \cite{Barthe07} was to close
this leak (formally called \textit{covert channel}) by annotating the byte-code
of the program\footnote{most mobile devices use some sort of byte-code e.g. Android} with security
labels according to the confidentiality of the information, which causes the scheduler to hide
threads that leak information from the attacker.

\section{Introduction to the research area}
\label{sec:introduction}
The challenge to protect the confidentiality of information processed by
a computer lead thirty years ago (c.f. \cite{Zdancewic04}) to the research
on the security of information-flows, because of the shortcomings of the standard methods.
These standard methods are: Access control and encryption. Where
access control ensures that the data is only accessed by an authorized entity,
and encryption ensures that data can be securely transmitted over an insecure channel.
But all these methods cover the release of information and not the further
propagation (i.e. how the information is used after having access to it or decrypting
it). This is the place where information-flow security comes into play.

The idea is to \textit{"track and regulate"}\cite{Zdancewic04} where information flows to
prevent the leak of secret information. The ways on which the information flows
through the system are called channels. Channels that are not indented to transport
information are called covered channels. These covered channels can
be classified into the following categories \cite{Sabelfeld03} (in the following examples $h$ is a secret
variable and $l$ a variable that is public):

\begin{itemize}
\item \textit{Explicit and implicit flows}: In an explicit flow the information is leaked
      directly into some public variable (e.g. $l := h$), while in implicit flows the information
      is leaked through the control structure of the program (e.g. $\text{if}\ h = 1\ \text{then}\ l := 1\ \text{else}\ l := 0$).
\item \textit{Probabilistic channels}: If an attacker is able to run a computation multiple times,
      he might be able to obtain information by looking at the probability distribution of the
      public outputs.
\item \textit{Power channels}: If the attacker has physical access to the computer or at least can figure out
      the power consumption, he might be able to obtain data through the changing power consumption.
\item \textit{Resource exhausting channels}: Information also may be leaked through the exhaustion
      of finite (mostly physical) resources (e.g. a buffer overflow).
\item \textit{Termination channels}: The attacker can obtain information through the termination
      or non-termination of the computation or program.
\item \textit{Timing channels:} Information can be obtained through the time at which this action
      occurs. On the one hand \textit{external} timing channels cover the actions like the termination
      of the program, where the attacker obtains information from the total execution time of the program.
      On the other hand \textit{internal} timing channels occur in multithreaded programs through the timing difference
      between threads. For example the two threaded (denoted with $||$) program \[\text{if}\ h = 1\ \{\text{sleep}(100)\}\ ;\ l := 1\ ||\ \text{sleep}(50)\ ;\ l := 0\]
      executed with most schedulers will leak the information of $h$ into $l$ through the timing difference
      between $h = 1$ and $h = 0$.
\end{itemize}

The notion of confidentiality that is used in most work in the area of information-flow security is \textit{noninterference}.
This policy is defined in various ways but the essence is, that it requires \textit{"that secret information
[do] not affect [the] publicly observable behavior of a system"}\cite{Zdancewic04}\footnote{Even if the attacker has full access to the
source code of the program.}. For most real world
applications this policy is far too strict, because it forbids useful programs like password checkers.\footnote{A
password checker needs to reveal whether the user input was correct or not.} Approaches that allow
controlled release of information are called declassification. The kinds of declassification are \textit{what}
information is released, \textit{who} released this information, \textit{where} in the system the information
is released and \textit{when} the information is released (c.f. \cite{Sabelfeld05}). For these approaches it is necessary that an
active attacker can only know as much as a passive attacker.

Mechanisms for controlling the information flow can be implemented either dynamic or static. Dynamic approaches\footnote{
The taint mode of the programming language Perl uses this mechanism.}
label the information with security labels and propagate these labels wherever the information is used. Static
approaches\footnote{This mechanism is implemented in Jif for Java and Flow Caml for Caml.} analyze the program code and are
therefor far more promising, because with these approaches it is possible to check all evaluation paths.

@@

\section{Summary of the article}
\label{sec:discussion}
In this summary I will follow mainly the structure of the original
paper \cite{Barthe07} which is as follows. First the basic
terms and definitions for multithreaded programs and the scheduler are
laid. After that the notion of security we want to achieve is presented.
Along with this a skeleton of a type system is described, which ensures
that a program typeable in this type system is secure.
The proof that this holds is sketched in the following section.
In the last section the (by now) abstract framework is instantiated with a
concrete example.

@@

These instructions are not further specified, but contain an instruction
to create a new thread (\texttt{start \textit{pc}} where pc is the start
instruction of the new thread). The instructions without \texttt{start}
are called $SeqIns$.

@@

The next thing introduced are the security levels. We assume the attacker
"is" a level $k$. From this assumption we can reduce every set of security levels
w.l.o.g into $\text{Level} = \{low, high\}$, where $low < high$, by mapping elements
that are no more sensitive than $k$ to $low$ and all other elements
-- including incomparable ones -- to $high$. It is also assumed, that access
control works correctly (i.e. the attacker can not access $high$ elements
directly).

@@

At this point a first simplification can be made. Looking a state $s \in \text{ConcState}$ we can
first extract the active threads (\texttt{s.act}) by taking the domain of the first
component. According to a security environment we can classify these threads,
with respect to their current program point (\texttt{s.pc(tid)} where $tid \in \text{Thread}$),
into \textit{low} threads if the current program point is low. In \textit{high}
threads iff the current program point is high. In \textit{always high} threads
if the current program point is always high and in \textit{hidden} threads, if
the current program point is high but not always high.

The latter two are the interesting ones. If a thread is \textit{always high} it
can not leak any information into low, because it never gets in touch
with low program points. So these threads can safely be interleaved between all
other threads by the scheduler.

@@

At this point no concrete scheduler is defined to make the framework applicable
for a wide class of schedulers. A scheduler is in this class if it can be
modeled as a function $pickt: \text{ConcState} \times \text{History} \rightharpoonup \text{Thread}$
which satisfies the following constraints:

\begin{enumerate}
\item It always picks an active thread
\item If there is a hidden thread, always choose high or always high threads\footnote{With this constraint the interleaving of always high threads is realized}
\item Only use low information and the low part of the history to choose a
      new thread
\end{enumerate}

To define what noninterference means we need a notion of execution. It is assumed
that we have a sequential execution relation $\leadsto_{seq} \subseteq \text{SeqState} \times
\text{SeqState}$. One step execution for the multithreaded language $\leadsto_{seq}
\subseteq (\text{ConcState} \times \text{History}) \times (\text{ConcState} \times \text{History})$ is defined
by the rules in figure \ref{fig:multithreaded-execution}.

@@

The scheduler is allowed to pick a new thread after every transition of $\leadsto_{seq}$. The
intuition of the first rule is, that the scheduler picks a new thread $coed$ which is at
program point $i$. This program point maps to an sequential instruction. The execution of
this instruction leads to a local state $\sigma$ and a global memory $\mu$. If this is the case
and the instruction $i$ was not the last instruction (i.e. $\sigma.pc \neq \mathtt{exit}$),
then the concurrent transition can be made. In this transition the concurrent state is updated
according to the results of the sequential execution and the thread id furthermore security level of $i$
is recorded in the history.

The second rule covers the case in which the thread picked by the scheduler has only the current
instruction before terminating (i.e. $\sigma.pc = \mathtt{exit}$). In the concurrent transition
the current thread id is removed from the concurrent state and everything else is like in the
first rule.

The third rule introduces the dynamic creation of new threads. In this rule the function $fresh_{l}$
\footnote{It is assumed that $fresh_l(tid) \neq fresh_{l'}(tid) \Leftrightarrow l \neq l'$.}
takes a set of thread identifiers and returns a new thread identifier at level $l$ and $\lambda_{init} : \mathcal{P}
\rightarrow \text{LocState}$ produces an initial state with the program pointer at the given program point.
If the current instruction at program point $i$ is the \texttt{start} instruction, a new thread
identifier is picked w.r.t. the security level of $i$, then program counter of the current thread
is increased by one, to step to the next instruction. The resulting concurrent state includes the
updated state of the current thread and the initial state for the new thread. The global memory
is not modified and the history is extended with the current thread id and the security level of $i$.

@@

The intuition is that the state $s$ evaluates according to a history $h$ to a final global memory $\mu$ iff there is a
sequence of concurrent executions that terminates (i.e. there are no active threads left) and has the
global memory $\mu$. $\leadsto_{conc}^*$ is the reflexive and transitive closure of of $\leadsto_{conc}$.

@@

Now we can define our goal: Noninterference. We define noninterference
in accordance to an indistinguishability relation $\sim_g$ on global memories.
Barthe et. al. state that it is not necessary -- for the purpose of the paper --
to specify the definition of this relation. But to get a feeling for this
relation, one can define it as: $\mu \sim_g \mu' \Leftrightarrow \mu|_{low} =
\mu'|_{low}$ where $\mu|_{low}$ projects out all high elements. Based on that
a program P is non-interfering if for all global memories $\mu_1, \mu_2, \mu_1'$
and $\mu_2'$ it holds that:

\begin{align*}
\mu_1 \sim_g \mu_2 \land (P,\mu_1 \Downarrow_{conc} \mu_1') \land (\text{P},\mu_2 \Downarrow_{conc} \mu_2') \Rightarrow \mu_1' \sim_g \mu_2'
\end{align*}

\subsection{Type system}
\label{sec:typesystem}
The type system is the core part of the framework in the sense that it
enforces the previously defined noninterference property. Thus every
program which is typeable, is non-interfering.

@@

The intuition of the typing judgment is, that if we execute the instruction at program
point $i$ w.r.t. the security environment $se$ and the current type is $\textsc{s}$, then the
type after the execution is $\textsc{t}$.

@@

A program is typeable in this type system (written $\mathcal{S}, se \vdash P$),
where $\mathcal{S}$ is a function $\mathcal{S}: \mathcal{P} \rightarrow \text{LType}$
that maps a local type to every program point
and $se$ a security environment, iff $\mathcal{S}$ maps every initial program point
\footnote{including the ones of spawned threads} to $\textsc{t}_{init}$ and that for every program point $j$, which is an
successor of $i$, there is a type $s \in \text{LType}$, such that $\textsc{s}$ is lower bounded
by $\mathcal{S}(j)$ and $se, i \vdash \mathcal{S}(i) \Rightarrow \textsc{s}$ holds.

@@

\begin{theorem}
If the scheduler is secure and $se, \mathcal{S} \vdash \text{P}$, then $\text{P}$ is non-interfering.
\end{theorem}

@@

The domain of next (i.e. NePd) captures the high, but not always high program points (i.e. the ones
that can result in a hidden thread). The property NeP1, states that two directly successive high
program points have the same program point in which the thread becomes visible again. The property NeP2
is the counterpart of the NeP1: If a low program point follows a high program point, this low program
point is the result of the next function for the high program point. NeP3 and NeP4 denote that the
next of an outermost if respective while instruction is at least after the control dependence region.

@@

where $op \in \{+,-,\times,/\}$, $n \in \mathbb{Z}$ and $x$ are variables. The
operational semantics are standard and not explicitly necessary for the following
instantiation, therefor they are omitted.

@@

Besides this concrete language we need to define a type system to enforce noninterference
according to section \ref{sec:typesystem}. The local types are defined by a stack of security levels
$\text{LType} = \text{Stack}(\text{Level})$ and $\textsc{t}_{init}$ with the empty stack. The typing rules defined in
figure~\ref{fig:multithreaded-typing-rules} are extended with rules for the concrete instructions
of the assembly language in figure~\ref{fig:transfer-rules}. In this summary I only explain two of them, the rest follows
in a similar manner.

@@

The security levels in $LType$ are the security levels of the operands in the local state. To
express our security policy, we declare a function $\Gamma(x)$ that assigns to every variable
$x$ a security level. If we now want to \texttt{store} the top of the stack into the variable
$x$, one premise is that the security level of the program point $se(i)$ \textit{and} the
security level of the value on top of the operand stack is \textit{lower or equal} to the
security level that was assigned by $\Gamma$ to the variable $x$. If this is the case the head
of $LType$ is removed. The $\sqcup$ operator is like a logical "or", it returns the higher of
the given security values.

@@

Now we can define $tregion$ according to a compilation function $\mathcal{C}$. I won't define
the compilation function here, because it is not essential how it exactly looks like.

@@

\section{Comparison with other approaches/further work}
\label{sec:furtherwork}
In this section I will first give a short overview of two other approaches and then
compare them with the approach presented before and draw a conclusion.

@@

They introduced the language $\lambda_{SEC}^{PAR}$ which supports higher-order functions, an
unbounded number of threads, synchronization (via join patterns), message passing and shared memory. Regardless of those realistic
features $\lambda_{SEC}^{PAR}$ is not intended to server as a user-lever programming language because its
syntax and type system is too awkward. Instead it is only used a model for studying information flow.

Low-security observational determinism is used to enforce noninterference. The idea is
to make the naturally nondeterministic system, deterministic from the point of view
of the attacker.

@@

\subsection{Flexible Scheduler-Independent Security\cite{Mantel10}}
This approach developed by Mantel and Sudbrock has the goal to be on the one hand
scheduler independent and on the other not too restrictive.

To reach this goal they model the scheduling explicit but general enough to apply
to a wide class of common schedulers. So rather than talking about program states,
they introduce system configurations which contains a list of threads (thread pool) the global
memory and the scheduler state. The scheduler is able to store information in its
state and can retrieve input (e.g. the number of active threads). Based on this
information the scheduler makes decisions. To cover nondeterministic schedulers
(e.g. the uniform scheduler) the decision making of the scheduler and the execution
of the program is probabilistic.

Based on this scheduler model, a schedule-specific security property $\mathcal{S}$ is
defined on thread pools. Intuitively a thread pool is $\mathcal{S}$-secure if the
probability of running a program with system configuration $conf$ under the scheduler $\mathcal{S}$
resulting in a global memory $m$ is the same as running it with configuration $conf'$ resulting in $m'$,
and satisfying that $m =_L m'$.

@@

In the end Barthe et. al. presented a framework with reaches many goals (like permissiveness, statically check-able
and requires no intervention from the programmer) but needs to modify the scheduler interface, which makes it
hard to use for real programs. In addition they did not how to implement the type-annotations in a real
language.
